Efficient and Effective Metric Space Indexing for Complete Record Linkage

### 2017-11-20 paper.tex
Update on Overleaf.

\begin{abstract}
%Probabilistic
Record linkage is the process of identifying records that refer to the
same real-world entities, in situations where entity identifiers are
unavailable. Records are linked on the basis of similarity between
common attributes, with every pair of records being classified as a
match or non-match depending on the degree of similarity between them.
Record linkage is usually performed in a three-step process, where first
groups of similar candidate records are identified using indexing. Only
pairs within the same group are then compared in more detail, and
finally classified.
%
Even state-of-the-art indexing techniques, such as Locality Sensitive
Hashing, have potential drawbacks. They may fail to group together some
true matching records with high similarity. Conversely, they may group
records with low similarity, leading to high computational overhead.
%
We propose using metric space indexing to perform \emph{complete} record
linkage, which results in a parameterless record linkage process
combining indexing, comparison and classification into a single step
delivering high effectiveness and efficiency.

@@

To overcome data quality issues such as typographical errors and
variations (which are common in name and address values~\cite{Chr12}),
approximate string comparison functions (such as edit distance, the
Jaro-Winkler comparator, or Jaccard similarity~\cite{Chr12}) are used to
compare pairs of records, leading to a vector of similarities (one
similarity per attribute/field compared) for each pair. These similarity
vectors are then used to classify the compared record pairs into matches
(where it is assumed both records in a pair correspond to the same
real-world entity) and non-matches (where the records are assumed to
correspond to two different entities). Various classification methods
have been employed in record linkage~\cite{Chr12,Don15}, ranging from
simple threshold based to sophisticated clustering, supervised
classification techniques, and active learning approaches~\cite{Wan15}.

@@

\smallskip
\textbf{Indexing for Record Linkage:}
Techniques to link records across datasets have been investigated for
over five decades~\cite{Fel69,New59} with the scalability of linking
being an ongoing challenge as datasets continue to grow in size and
complexity. Traditional blocking~\cite{Chr12b} uses a set of attributes
(known as a \emph{blocking key}) to insert records that share the same
value(s) in their blocking key into the same block. Only records within
the same block are  compared to each other. To overcome variations and
misspellings, the values used in blocking keys can be phonetically
encoded using functions such as Soundex, NYSIIS, or
Double-Metaphone~\cite{Chr12}. These techniques convert a string  into a
code according to its pronunciation, assigning the same code to similar
sounding names (such as `Gail' and `Gayle').

>>>>

\begin{abstract}
%Probabilistic
Record linkage is the process of identifying records that refer to the
same real-world entities, in situations where entity identifiers are
unavailable. Records are linked on the basis of similarity between
common attributes, with every pair of records being classified as a
link or non-link depending on the degree of similarity between them.
Record linkage is usually performed in a three-step process, where first
groups of similar candidate records are identified using indexing. Only
pairs within the same group are then compared in more detail, and
finally classified.
%
Even state-of-the-art indexing techniques, such as Locality Sensitive
Hashing, have potential drawbacks. They may fail to group together some
true matching records with high similarity. Conversely, they may group
records with low similarity, leading to high computational overhead.
%
We propose using metric space indexing to perform \emph{complete} record
linkage, which results in a parameterless record linkage process
combining indexing, comparison and classification into a single step
delivering high effectiveness and efficiency.

@@

To overcome data quality issues such as typographical errors and
variations (which are common in name and address values~\cite{Chr12}),
approximate string comparison functions (such as edit distance, the
Jaro-Winkler comparator, or Jaccard similarity~\cite{Chr12}) are used to
compare pairs of records, leading to a vector of similarities (one
similarity per attribute/field compared) for each pair. These similarity
vectors are then used to classify the compared record pairs into links
(where it is assumed both records in a pair correspond to the same
real-world entity) and non-links (where the records are assumed to
correspond to two different entities). Various classification methods
have been employed in record linkage~\cite{Chr12,Don15}, ranging from
simple threshold-based to sophisticated clustering, supervised
classification techniques, and active learning approaches~\cite{Wan15}.

@@

Techniques to link records across datasets have been investigated for
over five decades~\cite{Fel69,New59}, with the scalability of linking
being an ongoing challenge as datasets grow in size and
complexity. Traditional blocking~\cite{Chr12b} uses a set of attributes
(known as a \emph{blocking key}) to insert records that share the same
value(s) in their blocking key into the same block. Only records within
the same block are  compared to each other. To overcome variations and
misspellings, the values used in blocking keys can be phonetically
encoded using functions such as Soundex, NYSIIS, or
Double-Metaphone~\cite{Chr12}. These convert a string into a
code according to its pronunciation, assigning the same code to similar
sounding names (such as `Gail' and `Gayle'). Multiple blocking keys may
be used to deal with the problem of missing attribute values.
